# Introduction {#intro}

`r disclaimer()` 

`r button_code()`

## Diving into the CRAN & BioConductor database

Before going deeper into the licenses, let's first have a look into the licenses currently used on CRAN & BioConductor. 

```{r}
library(tidyverse)
theme_set(theme_minimal())
```

_Note: this version has been compiled on :_

```{r}
Sys.time()
```


### Get all the packages

First, let's get a list of all the packages. 

We'll extract the list of packages from CRAN using the `CRAN_package_db()` function from `{tools}`, and the ones from BioConductor by reading the file at `"https://bioconductor.org/packages/release/bioc/src/contrib/PACKAGES"`. These two datasets contain the metadata about the packages, as they are provdided in the DESCRIPTION file. 

```{r warning = FALSE}
# Getting the DB
cran_db <- tools::CRAN_package_db() %>% 
  thinkr::clean_names() %>%
  mutate(src = "cran")

bioc_db <- read.dcf(url("https://bioconductor.org/packages/release/bioc/src/contrib/PACKAGES")) %>% 
  as.data.frame() %>% 
  thinkr::clean_names()  %>%
  mutate(src = "bioc")

full_db <- bind_rows(cran_db, bioc_db)
```

These two lists contain: 

- `r nrow(cran_db)` packages from CRAN
- `r nrow(bioc_db)` packages from BioConductor

So a total of `r nrow(cran_db) + nrow(bioc_db)` packages.

### Data cleansing
 
Let's now select the columns of interest: the ones containing the informations about licenses, and have a look at the 6 first rows of the result. 

```{r}
db <- full_db %>% 
  select(package, contains("license"), src) 
head(db) 
```

## Extracting information

### Counting the licenses

```{r}
# Number of licences
n_licence <- db %>%
  distinct(license) %>% 
  count() %>% 
  pull(n)
```

In our dataset, we can find `r n_licence` different licences.

```{r}
# Number of licences with 'file'
db <- db %>% 
  mutate(has_file_license = str_detect(license, "file"))
n_has_file_license <- db %>% 
  count(has_file_license)

n_file <- n_has_file_license %>% 
  filter(has_file_license) %>% 
  pull(n)
```

Out of these `r nrow(db)` packages, `r n_file` contain a licence file — see [#TODO add ref]() for more info about adding licence files. 

### Packages with non-foss / restrictive-use licenses

The dataset also contains a column called `license_is_foss` and one called `License_restricts_use`.

> Fields ‘License_is_FOSS’ and ‘License_restricts_use’ may be added by repositories where information cannot be computed from the name of the license. ‘License_is_FOSS: yes’ is used for licenses which are known to be FOSS, and ‘License_restricts_use’ can have values ‘yes’ or ‘no’ if the LICENSE file is known to restrict users or usage, or known not to.

<div class = "right"> 
[Writing R Extensions](https://cran.r-project.org/doc/manuals/r-release/R-exts.html#Licensing)
</div>

```{r}
license_is_foss <- db %>% 
  count(license_is_foss) %>%
  drop_na()
```

The `license_is_foss` (Free Open Source Software) column is not that much informative, as only `r sum(license_is_foss$n)` packages have provided it. But what can be interesting is the `r license_is_foss %>% filter(license_is_foss == "no") %>% pull(n)` package(s) that contains a non-foss license: 

```{r}
full_db %>%
  filter(license_is_foss == "no") %>% 
  select(package, license, src)
```

We'll get back on this license in [#TODO add ref]().

```{r}
license_restricts_use <- db %>% 
  count(license_restricts_use) %>%
  drop_na()
```


Let's now have a look at the `license_restricts_use` column from the dataset. Few packages (`r nrow(license_restricts_use)`) have this field filled in their `DESCRIPTION`. Just as before, it can be interested to have a look at the packages that have do restrict the use. 

+ All the packages with a licence that restricts the use

```{r}
db %>% 
  filter(license_restricts_use == "yes") %>%
  select(package, license, src)
```

+ And if we want to have a look at the number of each of these licenses:

```{r}
db %>% 
  filter(license_restricts_use == "yes") %>%
  count(license, sort = TRUE)
```

As we can see, the majority of packages with a licence that restricts the use are based on a license file. 

```{r}
system.file("LICENSE", package = "rngwell19937") %>% 
  readLines() %>% 
  glue::as_glue()
```


```{r}
system.file("LICENSE", package = "gpclib") %>% 
  readLines() %>% 
  glue::as_glue()
```

We'll get back on this license in [#TODO add ref]().

## Getting a more global idea

There are several ways to insert a license : 

+ It can be restricted to precise version (e.g. `GPL-3`)
+ It can be "at least" a version (e.g. `GPL (>= 2)`)
+ It can be two licenses (e.g. `GPL-2 | GPL-3`)
+ It can be accompanied with a LICENSE file (e.g. `LGPL-3 | file LICENSE`)

Also, the license number can be either a number or a number + a zero (e.g `GPL (>= 2.0)`). So far, we have counted the number of different licenses based on a strict match of characters. So `GPL (>= 2.0)` and `GPL (>= 2)` are counted as two different licenses. 

In order to get a broader idea, we'll clean this column so that we can gather the licenses together. 

```{r}
cleaner <- compose(
  # Removing the file license elements
  ~ gsub("filelicense", "", .x),
  ~ gsub("filelicense", "", .x),
  # Removing empty spaces
  ~ gsub(" ", "", .x),
  # Removing all non-alpha numeric character
  proustr::pr_keep_only_alnum, 
  # Removing the .0 at the end of numbers
   ~ gsub("\\.0", "", .x), 
  tolower
)
db <- db %>%
 mutate(clean_license = cleaner(license)) 

db %>%
  count(clean_license, sort = TRUE)
```

As you can see, we now have `r length(unique(db$clean_license))`, while we originally had `r length(unique(db$license))` different licenses. 

_Note: this process was designed to have a broad overview of the licenses, and gathered together the `GPL (>= 2.0)` and `GPL-2` for example. It's designed more as a tool for general overview._

### How many GPL related licenses?

Let's have a look at how many licenses are GPL realted, by detecting the `"gpl"` string in the license column.

```{r}
# How many GPL based licenses ? 
n_gpl <- str_detect(db$clean_license, "gpl") %>%
  sum(na.rm = TRUE)
```

+ `r n_gpl` packages have a license which is a version of `GPL`
+ Which makes `r (n_gpl / nrow(db)) * 100`% of the packages.

### The 10 most frequent licenses?

Let's see which are the 10 most frequent licenses: 

```{r}
db %>%
  count(clean_license, sort = TRUE) %>%
  top_n(10)
```

## A visual representation of the licenses

```{r}
db %>%
  count(clean_license, sort = TRUE) %>%
  top_n(10, n) %>%
  ggplot(aes(reorder(clean_license, n), n )) + 
  geom_col() + 
  coord_flip() +
  labs(
    title = "10 most used licenses on the CRAN / Bioconductor",
    subtitle = glue::glue("Made on the {Sys.Date()}"), 
    x = "License", 
    y = "Volume"
  ) 
  

db %>%
  filter(src == "cran") %>% 
  count(clean_license, sort = TRUE) %>%
  top_n(10, n) %>%
  ggplot(aes(reorder(clean_license, n), n )) + 
  geom_col() + 
  coord_flip() +
  labs(
    title = "10 most used licenses on the CRAN",
    subtitle = glue::glue("Made on the {Sys.Date()}"), 
    x = "License", 
    y = "Volume"
  ) 

db %>%
  filter(src == "bioc") %>% 
  count(clean_license, sort = TRUE) %>%
  top_n(10, n) %>%
  ggplot(aes(reorder(clean_license, n), n )) + 
  geom_col() + 
  coord_flip() +
  labs(
    title = "10 most used licenses on Bioconductor",
    subtitle = glue::glue("Made on the {Sys.Date()}"), 
    x = "License", 
    y = "Volume"
  ) 

```

`r hide_code()`
