# Introduction {#intro}

## Diving into the CRAN & BioConductor database

```{r}
library(tidyverse)
```

- Getting all the packages 

```{r}
# Done on
Sys.time()
# Getting the DB
cran_db <- tools::CRAN_package_db() %>% thinkr::clean_names()

bioc_db <- read.dcf(url("https://bioconductor.org/packages/release/bioc/src/contrib/PACKAGES")) %>% as.data.frame() %>% thinkr::clean_names()
```

 
```{r}
cran_small <- cran_db %>% 
  select(package, contains("license")) %>%
  mutate(src = "cran")
bioc_small <- bioc_db %>% 
  select(package, contains("license")) %>%
  mutate(src = "bioc")
db <- rbind(cran_small, bioc_small)
```
 

How many different licenses?

```{r}
db %>%
  distinct(license) %>% 
  count()
```

Counting the number of package with a licence file:

```{r}
db$has_file_licence <- str_detect(db$license, "file")

# How many have a li
db %>% 
  count(has_file_licence)
```

Counting the `license_is_foss`:

```{r}
# Not that much information
db %>% 
  count(license_is_foss)
```

```{r}
db %>%
  filter(license_is_foss == "no")
```

How many package have a license with restrictive use?

```{r}
db %>% 
  count(license_restricts_use)

db %>% 
  filter(license_restricts_use == "yes") %>%
  select(package, license)

db %>% 
  filter(license_restricts_use == "yes") %>%
  count(license, sort = TRUE)
```

Cleaning the license name because to get a more global idea

```{r}
cleaner <- compose(
  # Removing the file licence elements
  ~ gsub("filelicense", "", .x),
  ~ gsub("filelicence", "", .x),
  # Removing empty spaces
  ~ gsub(" ", "", .x),
  # Removing all non-alpha numeric character
  proustr::pr_keep_only_alnum, 
  # Removing the .0 at the end of numbers
   ~ gsub("\\.0", "", .x), 
  tolower
)
db <- db %>%
 mutate(clean_license = cleaner(license)) 

db %>%
  count(clean_license, sort = TRUE)
```

How many GPL related licenses?

```{r}
# How many GPL based licenses ? 
n_gpl <- str_detect(db$clean_license, "gpl") %>%
  sum(na.rm = TRUE)
n_gpl

(n_gpl / nrow(db)) * 100
```

How may licenses are used just once? 

```{r}
db %>%
  count(clean_license) %>%
  filter(n == 1)
```

What are the 10 most frequent licenses?

```{r}
db %>%
  count(clean_license, sort = TRUE) %>%
  top_n(10)
```


```{r}
db %>%
  count(clean_license, sort = TRUE) %>%
  top_n(10) %>%
  ggplot(aes(reorder(clean_license, n), n )) + 
  geom_col() + 
  coord_flip() +
  labs(
    title = "10 most used licenses on the CRAN / Bioconductor",
    subtitle = glue::glue("Made on the {Sys.Date()}"), 
    x = "License", 
    y = "Volume"
  ) + 
  theme_minimal()

db %>%
  filter(src == "cran") %>% 
  count(clean_license, sort = TRUE) %>%
  top_n(10) %>%
  ggplot(aes(reorder(clean_license, n), n )) + 
  geom_col() + 
  coord_flip() +
  labs(
    title = "10 most used licenses on the CRAN",
    subtitle = glue::glue("Made on the {Sys.Date()}"), 
    x = "License", 
    y = "Volume"
  ) + 
  theme_minimal()

db %>%
  filter(src == "bioc") %>% 
  count(clean_license, sort = TRUE) %>%
  top_n(10) %>%
  ggplot(aes(reorder(clean_license, n), n )) + 
  geom_col() + 
  coord_flip() +
  labs(
    title = "10 most used licenses on Bioconductor",
    subtitle = glue::glue("Made on the {Sys.Date()}"), 
    x = "License", 
    y = "Volume"
  ) + 
  theme_minimal()

```
