# Introduction {#intro}

## Diving into the CRAN database

- Getting all the database 

```{r}
# Done on
Sys.time()
# Getting the DB
db <- tools::CRAN_package_db()
```
 
Doing a little bit of cleaning

```{r}
library(tidyverse)

# Cleaning the colnames 
db <- thinkr::clean_names(db)
```

How many different licenses?

```{r}
db %>%
  distinct(license) %>% 
  count()
```

Counting the number of package with a licence file:

```{r}
db$has_file_licence <- str_detect(db$license, "file")

# How many have a li
db %>% 
  count(has_file_licence)
```

Counting the `license_is_foss`:

```{r}
# Not that much information
db %>% 
  count(license_is_foss)
```

How many package have a license with restrictive use?

```{r}
db %>% 
  count(license_restricts_use)

db %>% 
  filter(license_restricts_use == "yes") %>%
  select(package, license)

db %>% 
  filter(license_restricts_use == "yes") %>%
  count(license, sort = TRUE)
```

Cleaning the license name because to get a more global idea

```{r}
cleaner <- compose(
  # Removing the file licence elements
  ~ gsub("filelicense", "", .x),
  ~ gsub("filelicence", "", .x),
  # Removing empty spaces
  ~ gsub(" ", "", .x),
  # Removing all non-alpha numeric character
  proustr::pr_keep_only_alnum, 
  # Removing the .0 at the end of numbers
   ~ gsub("\\.0", "", .x), 
  tolower
)
db <- db %>%
 mutate(clean_license = cleaner(license)) 

db %>%
  count(clean_license, sort = TRUE)
```

How many GPL related licenses?

```{r}
# How many GPL based licenses ? 
n_gpl <- str_detect(db$clean_license, "gpl") %>%
  sum()
n_gpl

(n_gpl / nrow(db)) * 100
```

How may licenses are used just once? 

```{r}
db %>%
  count(clean_license) %>%
  filter(n == 1)
```

What are the 10 most frequent licenses?

```{r}
db %>%
  count(clean_license, sort = TRUE) %>%
  top_n(10)
```


```{r}
db %>%
  count(clean_license, sort = TRUE) %>%
  top_n(10) %>%
  ggplot(aes(reorder(clean_license, n), n )) + 
  geom_col() + 
  coord_flip() +
  labs(
    title = "10 most used licenses on the CRAN",
    substile = glue::glue("Made on the {Sys.Date()}"), 
    x = "License", 
    y = "Volume"
  ) + 
  theme_minimal()
```

## Bioconductor 

> To Do

## GitHub

> Maybe search the top 1000 ? 
  

