# CRAN licenses

## "Officially Authorized" Licenses

From [R Licenses](https://www.r-project.org/Licenses/).

+ The “GNU Affero General Public License” version 3
+ The “Artistic License” version 2.0
+ The “BSD 2-clause License”
+ The “BSD 3-clause License”
+ The “GNU General Public License” version 2
+ The “GNU General Public License” version 3
+ The “GNU Library General Public License” version 2
+ The “GNU Lesser General Public License” version 2.1
+ The “GNU Lesser General Public License” version 3
+ The “MIT License”
+ The “Creative Commons Attribution-ShareAlike International License” version 4.0

## About R license 

```{r}
licence()
```

### GPL 2 & 3

#### GPL 2

- https://www.r-project.org/Licenses/GPL-2

#### GPL 3

- https://www.r-project.org/Licenses/GPL-3


### LGPL 2.1 or later

#### LGPL

- https://www.r-project.org/Licenses/LGPL-2

- https://www.r-project.org/Licenses/LGPL-2.1

- https://www.r-project.org/Licenses/LGPL-3

## The ten most used 

```{r}
db %>%
  count(license, sort = TRUE) %>%
  top_n(10)
```

#### AGPL

- https://www.r-project.org/Licenses/AGPL-3

### MIT

- https://www.r-project.org/Licenses/MIT

### CC0

- https://creativecommons.org/publicdomain/zero/1.0/

### Apache License

### BSD3 clause

- https://www.r-project.org/Licenses/BSD_3_clause

## Region-based licenses

Example: CeCILL - http://www.cecill.info/

```{r}
db$clean_license %>%
  grep(pattern = "cecill", x = ., value = TRUE) %>%
  length()
```

## Restrictive licenses

```{r}
db %>% 
  filter(license_restricts_use == "yes") %>%
  select(package, license)
```

- ACM 

- CC BY-NC-SA 4.0

- Artistic-2.0