# A Practical Guide {#practical}

`r disclaimer()`

So now that you've got a good grasp of the different licenses, how do you add these information to your package / R code?

## Filling the `DESCRIPTION` file

The most common way to share R code is to bundle it into a package. 
A package is a collection of R function, that comes with a `DESCRIPTION` file. 
This file contains the metadata about your package, which includes the license and the authors from the package.

### Adding licenses with `{usethis}`

If you are building a package, `{usethis}` comes handy when it comes to licenses: this package has a series of functions that help you add licenses, by adding the license name to the `DESCRIPTION`, but also by adding external file is necessary. 

For example, the MIT license needs to come with a file `LICENSE`, that contains the following:

```
YEAR:
COPYRIGHT HOLDER: 
```

You can add it with:

```{r eval = FALSE}
usethis::use_mit_license("NAME")
```

Note that if you want to add your name automatically, you can define `options(usethis.full_name = "My name")` in your `.Rprofile`. 

For example, here's mine:

```{r}
readLines("~/.Rprofile") %>%
  glue::as_glue()
```

Here are all the available licenses you can set with `{usethis}`:

```{r}
library(usethis)
 ls( "package:usethis", pattern = "license")
```

### Adding external contributors

If you are integrating someone else code inside your package, you need to state that in your package documentation. That's for example what is done in `{shiny}`:

```{r}
readLines(
  system.file("LICENSE", package = "shiny")
)[1:23] %>%
  glue::as_glue()
```

A good practice is also to include the author from these external libraries to your list of `Authors@R` in your `DESCRIPTION`, in the following form: 

```
person("Name", "Last Name", 
        role = c("ctb", "cph"),
        comment = "Author of library XXX")
```

Like what is done in `{shiny}`:

```{r}
read.dcf(
  system.file("DESCRIPTION", package = "shiny")
)[5] %>%
  glue::as_glue()
```

Other packages have a different, but still valid approach:
```{r}
read.dcf(
  system.file("DESCRIPTION", package = "tripack")
)[4] %>%
  glue::as_glue()
```

## Dependencies exploration

One thing you might want to do when building a package is analyse its dependencies and whether or not they might impact yours. 
To do that, here is a little suggestion, based on `{pkgnet}`.

1. Download the CRAN package database:

```{r}
pkgdb <- tools::CRAN_package_db()
```

2. Then, we'll use the `{pkgnet}` package to extract dependencies recursively.

```{r}
get_deps <- function(pkg){
  DependencyReporter$new()$set_package(pkg)$nodes$node
}

gol <- get_deps("golem")

extract_from_cran_db <- function(deps, pkgdb){
  pkgdb %>%
    as_tibble(.name_repair = "universal") %>%
    filter(Package %in% deps) 
}

deps <- extract_from_cran_db(gol, pkgdb)

deps %>%
  count(
    License
  )

deps %>%
  summarize(
    any(License_restricts_use, na.rm = TRUE)
  )

```
