# Appendix 

## The “GNU Affero General Public License” version 3

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("https://www.r-project.org/Licenses/AGPL-3"))
```

## The “Artistic License” version 2.0

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("https://www.r-project.org/Licenses/Artistic-2.0"))
```


## The “BSD 2-clause License”

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("https://www.r-project.org/Licenses/BSD_2_clause"))
```


## The “BSD 3-clause License”

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("https://www.r-project.org/Licenses/BSD_3_clause"))
```


## The “GNU General Public License” version 2

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("https://www.r-project.org/Licenses/GPL-2"))
```


## The “GNU General Public License” version 3

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("hhttps://www.r-project.org/Licenses/GPL-3"))
```


## The “GNU Library General Public License” version 2

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("https://www.r-project.org/Licenses/LGPL-2"))
```


## The “GNU Lesser General Public License” version 2.1

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("https://www.r-project.org/Licenses/LGPL-2.1"))
```


## The “GNU Lesser General Public License” version 3

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("https://www.r-project.org/Licenses/LGPL-3"))
```


## The “MIT License”

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("https://www.r-project.org/Licenses/MIT"))
```

## The “Creative Commons Attribution-ShareAlike International License” version 4.0

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("https://www.r-project.org/Licenses/CC-BY-SA-4.0"))
```
