# Appendix 

## The “GNU Affero General Public License” version 3 {#AGPL3}

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("https://www.r-project.org/Licenses/AGPL-3"))
```

## The “Artistic License” version 2.0 {#Artistic2}

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("https://www.r-project.org/Licenses/Artistic-2.0"))
```


## The “BSD 2-clause License” {#BSD2}

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("https://www.r-project.org/Licenses/BSD_2_clause"))
```


## The “BSD 3-clause License” {#BSD3}

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("https://www.r-project.org/Licenses/BSD_3_clause"))
```


## The “GNU General Public License” version 2 {#GPL2}

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("https://www.r-project.org/Licenses/GPL-2"))
```


## The “GNU General Public License” version 3 {#GPL3}

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("https://www.r-project.org/Licenses/GPL-3"))
```


## The “GNU Library General Public License” version 2 {#GLPL2}

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("https://www.r-project.org/Licenses/LGPL-2"))
```


## The “GNU Lesser General Public License” version 2.1  {#GLPL21}

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("https://www.r-project.org/Licenses/LGPL-2.1"))
```


## The “GNU Lesser General Public License” version 3 {#GLPL3}

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("https://www.r-project.org/Licenses/LGPL-3"))
```


## The “MIT License” {#MIT}

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("https://www.r-project.org/Licenses/MIT"))
```

## The “Creative Commons Attribution-ShareAlike International License” version 4.0 {#CCBYSA}

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("https://www.r-project.org/Licenses/CC-BY-SA-4.0"))
```

### GPL-compatible licenses {#GPLcompatible}

https://www.gnu.org/licenses/license-list.en.html#GPLCompatibleLicenses

+ GNU General Public License (GPL) version 3
+ GNU General Public License (GPL) version 2 
+ GNU Lesser General Public License (LGPL) version 3 
+ GNU Lesser General Public License (LGPL) version 2.1
+ GNU Affero General Public License (AGPL) version 3
+ GNU All-Permissive License 
+ Apache License, Version 2.0 
+ Artistic License 2.0 
+ Clarified Artistic License
+ Berkeley Database License
+ Boost Software License
+ Modified BSD license
+ CeCILL version 2
+ The Clear BSD License
+ Cryptix General License
+ eCos license version 2.0
+ Educational Community License 2.0
+ Eiffel Forum License, version 2
+ EU DataGrid Software License
+ Expat License
+ FreeBSD license 
+ Freetype Project License 
+ Historical Permission Notice and Disclaimer
+ License of the iMatix Standard Function Library
+ License of imlib2
+ Independent JPEG Group License 
+ Informal license
+ Intel Open Source License
+ ISC License
+ Mozilla Public License (MPL) version 2.0 
+ NCSA/University of Illinois Open Source License
+ License of Netscape JavaScript
+ OpenLDAP License, Version 2.7
+ License of Perl 5 and below
+ Public Domain 
+ License of Python 2.0.1, 2.1.1, and newer versions
+ License of Python 1.6a2 and earlier versions
+ License of Ruby 
+ SGI Free Software License B, version 2.0
+ Standard ML of New Jersey Copyright License
+ Unicode, Inc. License Agreement for Data Files and Software
+ Universal Permissive License (UPL) 
+ The Unlicense 
+ License of Vim, Version 6.1 or later 
+ W3C Software Notice and License 
+ License of WebM
+ WTFPL, Version 2
+ WxWidgets Library License 
+ X11 License 
+ XFree86 1.1 License
+ License of ZLib
+ Zope Public License, versions 2.0 and 2.1


