--- 
title: "Licensing R"
author: "Colin Fay"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
description: "Documenting current state of licensing in the R world."
---

# Licensing R 

Placeholder


## Disclaimer
## What is this book about 
## Special Thanks 

<!--chapter:end:index.Rmd-->


# Introduction {#intro}

Placeholder


## Diving into the CRAN & BioConductor database

<!--chapter:end:01-intro.Rmd-->


# What is a license (and why you should care)

Placeholder


## Definition 
### License
### Distribute
### Free software
### Copyleft
## Classification 
## Implication
## License compatibility

<!--chapter:end:02-what-is-license.Rmd-->


# Licenses in the real world

Placeholder


## R license 
### GPL 2 & 3
#### GPL 2
#### GPL 3
### LGPL 2.1 or later
#### LGPL
## Package Licenses
### "Officially Authorized" Licenses
### From Bioconductor 
### The ten most used in packages
#### MIT
### CC0
### Apache License 2
### BSD3 clause
## "Non-standard"
### Region-based licenses
## Restrictive licenses

<!--chapter:end:03-authorized.Rmd-->


# Licensing R code

Placeholder


## R-package 
### Code 
#### Defining your code reusability
#### Contribution 
### Dependancies
### Data Package
### Library wrappers
#### Including non R code (JavaScript, C++, CSS...)
## Documentation 
### Software documentation 
### Publication 

<!--chapter:end:04-r-license.Rmd-->

# Conclusion & Future work

## Current state of the licenses in the R world 

## Next steps

<!--chapter:end:05-summary.Rmd-->

`r if (knitr::is_html_output()) '
# References {-}
'`

<!--chapter:end:06-references.Rmd-->

# Appendix 

## The “GNU Affero General Public License” version 3

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("https://www.r-project.org/Licenses/AGPL-3"))
```

## The “Artistic License” version 2.0

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("https://www.r-project.org/Licenses/Artistic-2.0"))
```


## The “BSD 2-clause License”

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("https://www.r-project.org/Licenses/BSD_2_clause"))
```


## The “BSD 3-clause License”

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("https://www.r-project.org/Licenses/BSD_3_clause"))
```


## The “GNU General Public License” version 2

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("https://www.r-project.org/Licenses/GPL-2"))
```


## The “GNU General Public License” version 3

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("hhttps://www.r-project.org/Licenses/GPL-3"))
```


## The “GNU Library General Public License” version 2

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("https://www.r-project.org/Licenses/LGPL-2"))
```


## The “GNU Lesser General Public License” version 2.1

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("https://www.r-project.org/Licenses/LGPL-2.1"))
```


## The “GNU Lesser General Public License” version 3

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("https://www.r-project.org/Licenses/LGPL-3"))
```


## The “MIT License”

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("https://www.r-project.org/Licenses/MIT"))
```

## The “Creative Commons Attribution-ShareAlike International License” version 4.0

```{r result = "asis", echo = FALSE}
glue::as_glue(readLines("https://www.r-project.org/Licenses/CC-BY-SA-4.0"))
```

<!--chapter:end:07-appendix.Rmd-->

